---
title: "Bitcoin Momentum Feature Based Forecasting"
author: "Nico"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: no
editor_options:
  chunk_output_type: inline
---
# 1. Load Packages

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#",
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.align = "center",
  class.source = 'white',
  options(scipen=999) # suppress scientific notation
)
```

```{r load libraries}
# Load time series libraries
library(sweep)      # Broom-style tidiers for the forecast package
library(forecast)   # Forecasting models and predictions package
library(tidyquant)  # Loads tidyverse, financial pkgs, used to get data
library(timetk)     # Functions working with time series

# Library for scraping Bitcoin data
library(Quandl)

# Library for data manipulation
library(tidyverse)
```

# 2. Scrape Bitcoin Data

## 2.1 Quandl Functions

The quandl_tidy function is a wrapper around the Quandl function that returns a cleaner tibble.

```{r function for formatting imported data}
Quandl.api_key("5ydoG6gTCKjgzDpJp_1s") # 3GAtxPrAgoah7PyADPGy

quandl_tidy <- function(code, name) { 
  df <- Quandl(code) %>% 
    mutate(code = code, name = name) %>% 
    rename(date = Date, value = Value) %>% 
    arrange(date) %>% 
    as_tibble()
  return(df)
}
```

## 2.2 Bitcoin Exchange Rate Data

```{r import Bitcoin prices}
bitcoin_price <- Quandl("BCHARTS/BITSTAMPUSD") %>%
  arrange(Date) %>%
  as_tibble()

colnames(bitcoin_price) <- c("date", "open", "high", "low", "close", "volume_btc", "volume_currency", "weighted_price")
```


```{r remove local variables used to import data, echo = FALSE}
rm(quandl_tidy)
```

After importing the Bitcoin price data, we narrow our focus to the Bitcoin daily closing prices. Note that there are some days where the closing price is $0. In those cases, we will use the previous day's closing price to fill these values

```{r narrow scope to Bitcoin closing price}
bitcoin_price_tbl <- 
  bitcoin_price %>% 
  select(date, close) %>%
  mutate(close = ifelse(close == 0, NA, close)) %>% # replace 0s with NAs
  fill(close, .direction = c("down")) %>%           # fill NAs with previous day's closing price
  rename(price = close) %>%
  as_tibble()

head(bitcoin_price_tbl)
```

Split the data into training and testing data to test accuracy of forecasts. For now, we will set the training data period from January 2018 to June 2020, and the forecast period from July to November 2020.

```{r train test split}
# Keep last 365 days as test data
bitcoin_price_test_tbl <- 
  bitcoin_price_tbl %>%
  filter(date >= "2020-07-01" & date <= "2020-11-30")

# Remove test data rows from original dataset
bitcoin_price_train_tbl <- 
  bitcoin_price_tbl %>%
  filter(date >= "2018-01-01" & date <= "2020-06-30")
```



# 3. Auto-ARIMA

Sources: 

1. [Tidy Forecasting with sweep](https://www.business-science.io/code-tools/2017/10/25/demo_week_sweep.html)
2. [Amazon Stock Price Forecasting Using Time Series Analysis](https://rpubs.com/kapage/523169)

## 3.1 Visualisation

We begin by visualising the daily closing prices for Bitcoin over the time period of the data. 

```{r visualise Bitcoin daily closing prices}
# Plot daily Bitcoin closing prices
bitcoin_price_train_tbl %>%
    ggplot(aes(date, price)) +
    geom_line(col = palette_light()[6]) +
    geom_ma(ma_fun = SMA, n = 30, size = 1, col = palette_light()[2]) +
    theme_tq() +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    scale_y_continuous(labels = scales::dollar_format()) +
    labs(title = "Bitcoin Prices: Jan 2018 through Jun 2020")
```

## 3.2 ARIMA Forecasting

Convert Bitcoin daily closing price data from `tbl` to a `ts` object.

```{r convert from tbl to ts}
# Convert from tbl to ts
bitcoin_price_ts <- tk_ts(bitcoin_price_train_tbl, start = bitcoin_price_train_tbl$date[1], freq = 365)

# Check that ts-object has a timetk index
# This will be important when using sw_sweep() later
has_timetk_idx(bitcoin_price_ts)
```

Use the `auto.arima()` function from the `forecast` package to model the time series

```{r auto-arima}
# Model using auto.arima
fit_arima <- auto.arima(bitcoin_price_ts, lambda = "auto")

fit_arima
```

Tidy the model using `sweep` functions:

* `sw_tidy()`: Get model coefficients
* `sw_glance()`: Get model description and training set accuracy metrics
* `sw_augment()`: Get model residuals

```{r tidy model}
# sw_tidy - Get model coefficients
sw_tidy(fit_arima)

# sw_glance - Get model description and training set accuracy measures
sw_glance(fit_arima) %>%
    glimpse()

# sw_augment - get model residuals
sw_augment(fit_arima, timetk_idx = TRUE)

# plot residuals
sw_augment(fit_arima, timetk_idx = TRUE) %>%
    ggplot(aes(x = index, y = .resid)) +
    geom_point() + 
    geom_hline(yintercept = 0, color = "red") + 
    labs(title = "Residual Diagnostic") +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    theme_tq()
```

Use the `forecast()` function to forecast Bitcoin prices for the next month.

```{r construct forecast table}
# Forecast next 30 days
fcast_arima <- forecast(fit_arima, h = 335)

# Check if forecast has timetk index
has_timetk_idx(fcast_arima)

# tidy forecast output
fcast_tbl <- sw_sweep(fcast_arima, timetk_idx = TRUE)

fcast_tbl
```

Compare ARIMA forecasts with actual Bitcoin prices from Dec 2019 to Nov 2020.

```{r forecast visualisation}
# Visualize the forecast with ggplot
fcast_tbl %>%
    ggplot(aes(x = index, y = price, color = key)) +
    # 95% CI
    geom_ribbon(aes(ymin = lo.95, ymax = hi.95), 
                fill = "#D5DBFF", color = NA, size = 0) +
    # 80% CI
    geom_ribbon(aes(ymin = lo.80, ymax = hi.80, fill = key), 
                fill = "#596DD5", color = NA, size = 0, alpha = 0.8) +
    # Prediction
    geom_line() +
    # geom_point() +
    # Actuals
    geom_line(aes(x = date, y = price), color = palette_light()[[1]], data = bitcoin_price_test_tbl) +
    # geom_point(aes(x = date, y = price), color = palette_light()[[1]], data = bitcoin_price_test_tbl) +
    # Aesthetics
    labs(title = "Bitcoin Prices Forecast", x = "", y = "Closing Price",
         subtitle = "ARIMA(1,1,1) without Drift") +
    scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
    scale_y_continuous(labels = scales::dollar_format()) +
    scale_color_tq() +
    scale_fill_tq() +
    theme_tq()
```

Calculate errors for forecast Bitcoin prices.

```{r forecast error metrics}
# Investigate test error 
error_tbl <- left_join(bitcoin_price_test_tbl, fcast_tbl, by = c("date" = "index")) %>%
    rename(actual = price.x, pred = price.y) %>%
    select(date, actual, pred) %>%
    mutate(
        error     = actual - pred,
        error_pct = error / actual
        )

# Calculate test error metrics
test_residuals <- error_tbl$error
test_error_pct <- error_tbl$error_pct * 100 # Percentage error

me   <- mean(test_residuals, na.rm=TRUE)
rmse <- mean(test_residuals^2, na.rm=TRUE)^0.5
mae  <- mean(abs(test_residuals), na.rm=TRUE)
mape <- mean(abs(test_error_pct), na.rm=TRUE)
mpe  <- mean(test_error_pct, na.rm=TRUE)

tibble(me, rmse, mae, mape, mpe) %>% glimpse()
```


# 4. Time Series with Modeltime

The `modeltime` package allows us to easily build several time series models using the tidymodels framework and compare the predictions of all models on the same plot.

```{r modeltime forecast}
library(tidymodels)
library(modeltime)
library(lubridate)

# Model 1: auto_arima
model_fit_arima_no_boost <- arima_reg() %>%
    set_engine(engine = "auto_arima") %>%
    fit(price ~ date, data = bitcoin_price_train_tbl)

# Model 2: arima with XGBoost
model_fit_arima_boosted <- arima_boost(
    min_n = 2,
    learn_rate = 0.015
) %>%
    set_engine(engine = "auto_arima_xgboost") %>%
    fit(price ~ date + as.numeric(date) + factor(day(date), ordered = F),
        data =  bitcoin_price_train_tbl)

# Model 3: error-trend-season (ets)
model_fit_ets <- exp_smoothing() %>%
    set_engine(engine = "ets") %>%
    fit(price ~ date, data = bitcoin_price_train_tbl)

# Model 4: prophet
model_fit_prophet <- prophet_reg() %>%
    set_engine(engine = "prophet") %>%
    fit(price ~ date, data = bitcoin_price_train_tbl)

# Model 5: prophet with XGBoost
model_fit_prophet_boosted <- prophet_boost(
        seasonality_daily  = FALSE, 
        seasonality_weekly = FALSE, 
        seasonality_yearly = FALSE,
        changepoint_range  = 0.90,
        trees              = 300,
        mtry               = 0.50,
        min_n              = 30,
        learn_rate         = 0.15
    ) %>% 
    set_engine("prophet_xgboost") %>%
    fit(price ~ date, data = bitcoin_price_train_tbl)

# Add fitted models to a Modeltime Table
models_tbl <- modeltime_table(
    model_fit_arima_no_boost,
    model_fit_arima_boosted,
    model_fit_ets,
    model_fit_prophet,
    model_fit_prophet_boosted
)

# Calibrate the model to test data
calibration_tbl <- models_tbl %>%
    modeltime_calibrate(new_data = bitcoin_price_test_tbl)

# Visualise forecasted values for all models
calibration_tbl %>%
    modeltime_forecast(
        new_data    = bitcoin_price_test_tbl,
        actual_data = rbind(bitcoin_price_train_tbl,
                            bitcoin_price_test_tbl)
    ) %>%
    plot_modeltime_forecast(
      .legend_max_width = 25, # For mobile screens
      .interactive      = TRUE
    )
```

From the chart above, the predicted values generated by the normal `prophet` model stands out the most due to its variability as compared to the straight line predictions for the rest of the models.

```{r modeltime accuracy}
calibration_tbl %>%
    modeltime_accuracy() %>%
    table_modeltime_accuracy(
        .interactive = TRUE
    )
```

Comparing the accuracy of the 5 time series models we trained, the `prophet` models have the lowest MAPE scores. However, the prices of Bitcoin are unlikely to exhibit a steady rise as predicted by the  `prophet_boost()` model. Thus, we will treat the base `prophet` model (model 4) as the most ideal time series model.

```{r}
# Keep prophet forecasts
prophet_forecasts <-
    calibration_tbl %>%
    modeltime_forecast(
        new_data    = bitcoin_price_test_tbl,
        actual_data = rbind(bitcoin_price_train_tbl,
                            bitcoin_price_test_tbl)
    ) %>% filter(.model_id == 4) %>%
    select(.index, .value) %>%
    rename(date = .index, price = .value)

# Append the last row of train data to prospect forecasts
prophet_forecasts <-
    rbind(tail(bitcoin_price_train_tbl, 1),
          prophet_forecasts)

# Compute daily returns using tq_mutate
prophet_forecasts <- 
    prophet_forecasts %>%
    tq_mutate(select = price,
              mutate_fun = periodReturn,
              period = 'daily',
              type = 'arithmetic',
              col_rename = 'future_return')

# New column that contains 1 if Bitcoin prices increase, 0 otherwise
prophet_forecasts_sign <-
    prophet_forecasts %>%
    mutate(future_return_sign = ifelse(prophet_forecasts$price > lag(prophet_forecasts$price), 
                                       1, 0)
    ) %>%
    filter(date >= "2020-07-01")

prophet_forecasts_sign
```

