---
title: "Bitcoin Momentum Feature Based Forecasting"
author: "Nico"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: yes
      smooth_scroll: no
editor_options:
  chunk_output_type: console
---
# 1. Load Packages

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#",
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  fig.align = "center",
  class.source = 'white'
)
```

```{r}
library(tidyverse)
library(stringr)
library(gtrendsR)
library(Quandl)
library(quantmod)
library(RcppRoll)
library(lubridate)
library(tidyquant)
library(tidymodels)
library(tsfeatures)
library(slider)
library(timetk)
library(data.table)
```

# 2. Scrape Bitcoin Data

## 2.1 Quandl Functions

The quandl_tidy function is a wrapper around the Quandl function that returns a cleaner tibble.

```{r}
Quandl.api_key("5ydoG6gTCKjgzDpJp_1s") # 3GAtxPrAgoah7PyADPGy

quandl_tidy <- function(code, name) { 
  df <- Quandl(code) %>% 
    mutate(code = code, name = name) %>% 
    rename(date = Date, value = Value) %>% 
    arrange(date) %>% 
    as_tibble()
  return(df)
}
```

## 2.2 Bitcoin Exchange Rate Data

```{r}
bitcoin_price <- Quandl("BCHARTS/BITSTAMPUSD") %>%
  arrange(Date) %>%
  as_tibble()

colnames(bitcoin_price) <- c("date", "open", "high", "low", "close", "volume_btc", "volume_currency", "weighted_price")
```

## 2.3 Bitcoin Indicators

Data about bitcoin activity, transaction fees and mining.

```{r}
code_list <- list(c("BCHAIN/TOTBC", "Total Bitcoins"), 
                  c("BCHAIN/MKTCP", "Bitcoin Market Capitalization"), 
                  c("BCHAIN/NADDU", "Bitcoin Number of Unique Addresses Used"), 
                  c("BCHAIN/ETRAV", "Bitcoin Estimated Transaction Volume BTC"), 
                  c("BCHAIN/ETRVU", "Bitcoin Estimated Transaction Volume USD"), 
                  c("BCHAIN/TRVOU", "Bitcoin USD Exchange Trade Volume"), 
                  c("BCHAIN/NTRAN", "Bitcoin Number of Transactions"), 
                  c("BCHAIN/NTRAT", "Bitcoin Total Number of Transactions"), 
                  c("BCHAIN/NTREP", "Bitcoin Number of Transactions Excluding Popular Addresses"), 
                  c("BCHAIN/NTRBL", "Bitcoin Number of Tansaction per Block"), 
                  c("BCHAIN/ATRCT", "Bitcoin Median Transaction Confirmation Time"), 
                  c("BCHAIN/TRFEE", "Bitcoin Total Transaction Fees"), 
                  c("BCHAIN/TRFUS", "Bitcoin Total Transaction Fees USD"), 
                  c("BCHAIN/CPTRA", "Bitcoin Cost Per Transaction"), 
                  c("BCHAIN/CPTRV", "Bitcoin Cost % of Transaction Volume"), 
                  c("BCHAIN/BLCHS", "Bitcoin api.blockchain Size"), 
                  c("BCHAIN/AVBLS", "Bitcoin Average Block Size"), 
                  c("BCHAIN/TOUTV", "Bitcoin Total Output Volume"), 
                  c("BCHAIN/HRATE", "Bitcoin Hash Rate"), 
                  c("BCHAIN/MIREV", "Bitcoin Miners Revenue"), 
                  c("BCHAIN/BCDDE", "Bitcoin Days Destroyed"), 
                  c("BCHAIN/BCDDW", "Bitcoin Days Destroyed Minimum Age 1 Week"), 
                  c("BCHAIN/BCDDM", "Bitcoin Days Destroyed Minimum Age 1 Month"), 
                  c("BCHAIN/BCDDY", "Bitcoin Days Destroyed Minimum Age 1 Year") ,
                  c("BCHAIN/BCDDC", "Bitcoin Days Destroyed Cumulative"))

bitcoin_data <- tibble()

for (i in seq_along(code_list)) { 
  
  print(str_c("Downloading data for ", code_list[[i]][1], "."))
  
  bitcoin_data <- bind_rows(bitcoin_data, 
                            quandl_tidy(code_list[[i]][1], code_list[[i]][2]))
  
}

bitcoin_data <- bitcoin_data %>%
  select(-name) %>%
  spread(code, value)

colnames(bitcoin_data) <- make.names(colnames(bitcoin_data))
```

```{r, echo = FALSE}
rm(code_list, i, quandl_tidy)
```

# 3. Scrape Google Data

## 3.1 Create Date Set

```{r}
download_all <- FALSE

if (download_all == TRUE) {
  
  dates <- tibble(dates = ymd("2011-01-01") + months(0:120)) %>% filter(dates <= Sys.Date())

} else {
  
  dates <- tibble(dates = ymd("2017-01-01") + months(0:120)) %>% filter(dates <= Sys.Date())
  
}
```

## 3.2 Google Trends Function

```{r}
google_trends <- function(query, begin_date, end_date) {
  
  df <- gtrends(keyword = 'bitcoin',
                time = str_c(begin_date, ' ', end_date))[['interest_over_time']] %>%
    select(date, hits) %>%
    mutate(date = as.Date(date)) %>%
    as_tibble()
  
  return(df)
  
}
```

### 3.2.1 Google Trends Daily

```{r}
for (i in 1:nrow(dates)) {
  
  month <- dates[["dates"]][i]
  begin_date <- as.Date(month)
  end_date <- as.Date(month) + months(1) - days(1)
  end_date <- as.Date(ifelse(end_date >= Sys.Date(), Sys.Date(), end_date))

  df <- google_trends("bitcoin", begin_date, end_date)
  
  write_csv(df, file.path('pub_trends', str_c("google-trends-daily-", begin_date, "-", end_date, ".csv")))
  
}
```

### 3.2.2 Google Trends Monthly

```{r}
monthly <- google_trends('bitcoin', '2011-01-01', Sys.Date()) %>%
  rename(hits_monthly = hits)
```

### 3.2.3 Rebase

```{r}
bitcoin_google <- list.files('pub_trends') %>%
  map_df(~ read_csv(file.path('pub_trends', str_c(.)), col_types = c('Di'))) %>%
  rename(hits_daily = hits) %>%
  left_join(monthly) %>%
  fill(hits_monthly) %>%
  mutate(hits_monthly = as.numeric(hits_monthly),
         hits_daily = hits_daily * hits_monthly / 100)
```

```{r, echo = FALSE}
rm(download_all, dates, google_trends, monthly)
```

# 4. Scrape Other Data

```{r}
get_yahoo <- function(ticker) {
  
  df <- getSymbols(ticker, src = 'yahoo', auto.assign = FALSE, from = '1900-01-01')
  
  df <- df %>%
    as_tibble() %>%
    mutate(date = index(df))
  
  colnames(df) <- c("open", "high", "low", "close", "volume", "adjusted_close", "date", "ticker")
  
  return(df)
  
}

vix <- get_yahoo('^VIX') %>%
  select(date, vix = adjusted_close)
```

```{r, echo = FALSE}
rm(get_yahoo)
```

# 5. Feature Engineering

## 5.1 Cleaning Bitcoin Prices

Data errors are cleaned by using last observations carried forward.

```{r}
bitcoin_price[bitcoin_price == 0] <- NA
bitcoin_price <- bitcoin_price %>%
  map_df(na.locf)
```

## 5.2 Define Target

```{r}
bitcoin_model <- bitcoin_price %>%
  tq_mutate(select = close,
            mutate_fun = periodReturn,
            period = 'daily',
            type = 'arithmetic',
            col_rename = 'future_return') %>%
  mutate(future_return_sign = as.factor(ifelse(future_return > 0, 1, 0)),
         close = lag(close, 1),
         date = date - days(1)) %>%
  select(date, close, future_return, future_return_sign) 

bitcoin_model <- bitcoin_model[-1, ]
```

## 5.3 Bitcoin Price Change

```{r}
close_change <- shift(bitcoin_model$close, n = 1:90, type = 'lag', give.names = TRUE) %>% 
  as.data.frame() %>% 
  rename_all(funs(paste0(gsub('V1_lag_', 'close_change_', x = .), 'd')))

bitcoin_model <- bitcoin_model %>%
  cbind(close_change) %>%
  mutate_if(grepl('close_change_', names(.)), ~ close / . - 1)
```

```{r}
ggplot(bitcoin_model, aes(x = close_change_2d, fill = future_return_sign)) + 
  geom_histogram(binwidth = 0.0025, alpha = 0.5, position = "identity") + 
  coord_cartesian(xlim = c(-0.20, 0.20))
```
```{r}
ggplot(bitcoin_model, aes(x = close_change_2d, y = future_return)) + 
  geom_point(alpha = 0.25, colour = "blue") + 
  coord_cartesian(xlim = c(-0.20, 0.20), ylim = c(-0.10, 0.10)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_smooth(colour = "red")
```
```{r}
ggplot(bitcoin_model, aes(x = close_change_16d, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 0.01, alpha = 0.5, position = "identity") + 
  coord_cartesian(xlim = c(-0.50, 0.50))
```
```{r}
ggplot(bitcoin_model, aes(x = close_change_16d, y = future_return)) + 
  geom_point(alpha = 0.25, colour = "blue") + 
  coord_cartesian(xlim = c(-0.50, 0.50), ylim = c(-0.10, 0.10)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_smooth(colour = "red")
```
```{r}
ggplot(bitcoin_model, aes(x = close_change_57d, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 0.01, alpha = 0.5, position = "identity") + 
  coord_cartesian(xlim = c(-0.80, 0.80))
```
```{r}
ggplot(bitcoin_model, aes(x = close_change_57d, y = future_return)) + 
  geom_point(alpha = 0.25, colour = "blue") + 
  coord_cartesian(xlim = c(-0.50, 0.50), ylim = c(-0.10, 0.10)) + 
  geom_hline(yintercept = 0) + 
  geom_vline(xintercept = 0) + 
  geom_smooth(colour = "red")
```

## 5.4 Bitcoin Drawdown

```{r}
bitcoin_model <- bitcoin_model %>%
  mutate(close_drawdown = -1 * (1 - close / cummax(close)))
```

```{r}
ggplot(bitcoin_model, aes(x = close_drawdown, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 0.01, alpha = 0.5, position = "identity")
```

## 5.5 Rolling Daily Return Volatility

```{r}
close_sd <- frollapply(bitcoin_model$close_change_1d, 1:90, sd) %>%
  as.data.frame()
names(close_sd) <- paste0(sprintf('close_sd_%s', seq(1:90)), 'd')

bitcoin_model <- bitcoin_model %>%
  cbind(close_sd) %>%
  mutate(close_sd_1d = 0)
```

```{r}
ggplot(bitcoin_model, aes(x = close_sd_10d, fill = factor(future_return_sign))) +
  geom_histogram(binwidth = 0.001, position = "identity", alpha = 0.5)
```

```{r}
ggplot(bitcoin_model, aes(x = close_sd_90d, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 0.001, position = "identity", alpha = 0.5)
```

## 5.6 Number of Positive Days

```{r}
bitcoin_model <- bitcoin_model %>%
  mutate(close_positive = ifelse(close_change_1d > 0, 1, 0), 
         close_negative = ifelse(close_change_1d <= 0, 1, 0))

close_positive <- frollsum(bitcoin_model$close_positive, 1:90, align = 'right') %>%
  as.data.frame()
names(close_positive) <- paste0(sprintf('close_positive_%s', seq(1:90)), 'd')

bitcoin_model <- bitcoin_model %>%
  cbind(close_positive)
```

## 5.7 Number of Consecutive Positive and Negative Days

```{r}
bitcoin_model <- bitcoin_model %>% 
  mutate(close_positive_streak = close_positive * unlist(map(rle(close_positive)[["lengths"]], seq_len)), 
         close_negative_streak = close_negative * unlist(map(rle(close_negative)[["lengths"]], seq_len)))
```

```{r}
ggplot(bitcoin_model, aes(x = close_positive_streak, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.5)
```

```{r}
ggplot(bitcoin_model, aes(x = close_negative_streak, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.5)
```

```{r}
ggplot(bitcoin_model, aes(x = close_positive_26d, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.5)
```

```{r}
ggplot(bitcoin_model, aes(x = close_positive_63d, fill = factor(future_return_sign))) + 
  geom_histogram(binwidth = 1, position = "identity", alpha = 0.5)
```

## 5.8 google Trend Change

```{r}
bitcoin_model <- bitcoin_model %>% 
  left_join(bitcoin_google, by = 'date')

hits_daily <- shift(bitcoin_model$hits_daily, n = 1:30, type = 'lag', give.names = TRUE) %>% 
  as.data.frame() %>% 
  rename_all(funs(paste0(gsub('V1_lag_', 'hits_daily_', x = .), 'd')))

bitcoin_model <- bitcoin_model %>%
  cbind(hits_daily) %>%
  mutate_if(grepl('hits_daily_', names(.)), ~ hits_daily / . - 1)
```

```{r}
ggplot(bitcoin_model, aes(x = date)) + 
  geom_line(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_line(aes(y = hits_daily * 35), colour = "red", alpha = 0.6)
```

```{r}
ggplot(bitcoin_model %>% filter(date >= "2017-01-01"), aes(x = date)) + 
  geom_line(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_line(aes(y = hits_daily * 47), colour = "red", alpha = 0.6) + 
  geom_point(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_point(aes(y = hits_daily * 47), colour = "red", alpha = 0.6)
```

```{r}
ggplot(bitcoin_model %>% filter(date >= "2016-01-01", date <= "2017-01-01"), aes(x = date )) + 
  geom_line(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_line(aes(y = hits_daily * 47), colour = "red", alpha = 0.6) + 
  geom_point(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_point(aes(y = hits_daily * 47), colour = "red", alpha = 0.6)
```

```{r}
ggplot(bitcoin_model, aes(x = hits_daily_1d, y = future_return)) + 
  geom_point(alpha = 0.2)
```

```{r}
ggplot(bitcoin_model, aes(x = hits_daily_1d, fill = factor(future_return_sign))) + 
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.01) + 
  coord_cartesian(xlim = c(-0.50, 0.50))
```

### 5.8.1 Smoothed Google Trends

```{r}
hits_daily_sma <- frollmean(bitcoin_model$hits_daily, 1:10, align = 'right') %>%
  as.data.frame()
names(hits_daily_sma) <- paste0(sprintf('hits_daily_sma_%s', seq(1:10)), 'd')

bitcoin_model <- bitcoin_model %>%
  cbind(hits_daily_sma)
```

```{r}
ggplot(bitcoin_model, aes(x = date)) + 
  geom_line(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_line(aes(y = hits_daily_sma_5d * 47), colour = "red", alpha = 0.8)
```
```{r}
ggplot(bitcoin_model %>% filter(date >= "2017-01-01"), aes(x = date )) + 
  geom_line(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_line(aes(y = hits_daily_sma_5d * 47), colour = "red", alpha = 0.6) + 
  geom_point(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_point(aes(y = hits_daily_sma_5d * 47), colour = "red", alpha = 0.6)
```

```{r}
ggplot(bitcoin_model %>% filter(date >= "2016-01-01", date <= "2017-01-01"), aes(x = date )) + 
  geom_line(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_line(aes(y = hits_daily_sma_5d * 47), colour = "red", alpha = 0.6) + 
  geom_point(aes(y = close), colour = "blue", alpha = 0.8) + 
  geom_point(aes(y = hits_daily_sma_5d * 47), colour = "red", alpha = 0.6)
```

### 5.8.2 Google Trends Drawdown

```{r}
bitcoin_model <- bitcoin_model %>%
  mutate(hits_daily_drawdown = -1 * (1 - hits_daily / cummax(fill(bitcoin_model, hits_daily, .direction = "downup")$hits_daily))
  )
```

```{r}
ggplot(bitcoin_model, aes(x = hits_daily_drawdown, fill = factor(future_return_sign))) + 
  geom_histogram(position = "identity", alpha = 0.5, binwidth = 0.01)
```

## 5.9 VIX

```{r}
bitcoin_model <- bitcoin_model %>% 
  left_join(vix)
```

```{r}
ggplot(bitcoin_model, aes(x = date)) + 
  geom_line(aes(y = close), colour = "blue") + 
  geom_line(aes(y = vix * 30), colour = "red")
```

```{r}
ggplot(bitcoin_model %>% filter(date >= "2017-01-01"), aes(x = date)) + 
  geom_line(aes(y = close), colour = "blue") + 
  geom_line(aes(y = vix * 150), colour = "red")
```

## 5.10 Time Series Features

```{r}
ts_feature_set <- c(
    "entropy",         # Measures the "forecastability" of a series - low values = high sig-to-noise, large vals = difficult to forecast
    "stability",       # Means/variances are computed for all tiled windows - stability is the variance of the means
    "lumpiness",       # Lumpiness is the variance of the variances
    "max_level_shift", # Finds the largest mean shift between two consecutive windows (returns two values, size of shift and time index of shift)
    "max_var_shift",   # Finds the max variance shift between two consecutive windows (returns two values, size of shift and time index of shift)
    "max_kl_shift",    # Finds the largest shift in the Kulback-Leibler divergence between two consecutive windows (returns two values, size of shift and time index of shift)
    "crossing_points"  # Number of times a series crosses the mean line
)

bitcoin_model <- bitcoin_model %>%
  mutate(ts_features = slide(
    .x = future_return,
    .f = ~ tsfeatures(.x, features = ts_feature_set),
    .before = 90,
    .complete = TRUE)) %>% 
  unnest(ts_features)
```

## 5.11 Cleaning Data

```{r}
bitcoin_model <- bitcoin_model %>%
  filter(date >= '2013-10-01')

bitcoin_model <- bitcoin_model %>%
  select(-matches('close_sd_'),
         -matches('hits_daily_'),
         -matches('hits_daily_drawdown'))
```

# 6. Predictive Modeling

## 6.1 Recipe

```{r}
recipe_spec <- recipe(future_return_sign ~ ., data = bitcoin_model) %>%
  update_role(date, future_return, close, new_role = "ID")

rmarkdown::paged_table(recipe_spec %>% prep() %>% juice() %>% head())
```

## 6.2 XGBoost

### 6.2.1 XGBoost Model

```{r}
xgb_model <- boost_tree(learn_rate = 0.01,
                        tree_depth = 1,
                        min_n = 1,
                        mtry = 500,
                        trees = tune(),
                        stop_iter = 50) %>%
  set_engine('xgboost') %>%
  set_mode('classification')

xgb_model
```

```{r}
xgb_wflw <- workflow() %>%
  add_recipe(recipe_spec) %>%
  add_model(xgb_model)

xgb_wflw
```

### 6.2.2 Cross Validation

```{r}
resamples_cv <- recipe_spec %>%
  prep() %>%
  juice() %>%
  time_series_cv(
    date_var = date,
    initial = '3 month',
    assess = '3 month',
    skip = '3 month',
    cumulative = TRUE
  )
```

### 6.2.3 Building Model

```{r}
library(doParallel)
registerDoParallel()

xgb_params <- grid_max_entropy(trees(), size = 10)

xgb_model_trained <- tune_grid(
  xgb_wflw, 
  grid = xgb_params,
  metrics = metric_set(accuracy, roc_auc, mn_log_loss),
  resamples = resamples_cv,
  control = control_resamples(verbose = FALSE,
                              save_pred = TRUE,
                              allow_par = TRUE))

xgb_model_trained %>% collect_metrics()
```

### 6.2.4 Making Predictions

```{r}
best_params <- xgb_model_trained %>%
  select_best('mn_log_loss', maximise = FALSE)

xgb_model_best <- xgb_model %>%
  finalize_model(best_params)

xgb_wflw_best <- workflow() %>%
  add_recipe(recipe_spec) %>%
  add_model(xgb_model_best)

resampling_results <- xgb_wflw_best %>%
  fit_resamples(
    resamples = resamples_cv,
    control   = control_resamples(
      verbose   = FALSE, 
      save_pred = TRUE,
      allow_par = TRUE))

resample_predictions <- resampling_results %>%
  mutate(.testing = map(splits, testing)) %>%
  select(id, .predictions, .testing) %>%
  unnest(c(.predictions, .testing), names_repair = "universal") 
```

```{r}
signal <- resample_predictions %>%
  select(- c(future_return_sign...208, .pred_0)) %>%
  rename(future_return_pred = .pred_1,
         future_return_sign = future_return_sign...6) %>%
  arrange(date) %>%
  mutate(signal = ifelse(lag(future_return_pred, n = 1, default = 1) > 0.50, 1, 0),
         trading_cost = abs(signal - lag(signal, n = 1, default = 0)) * 0.003,
         return_buyhold = cumprod(1 + close_change_1d),
         return_model = cumprod(1 + close_change_1d * signal - trading_cost))
```

# 7 Evaluating Model

## 7.1 Visualising Signals

Predicted probability ranges from 0.428 to 0.577.

```{r}
ggplot(signal, aes(x = future_return_pred)) + 
  geom_histogram(binwidth = 0.001)
```

The following plots show the signal overlayed on top of the bitcoin price series. Dark blue indicates days where the model predicts a high probability of positive price movement and dark green indicates day where the model predicts a low probability of positive price movement. Purple-ish colors represent intermediate probabilities.

```{r}
ggplot(signal, aes(x = date, y = close, colour = future_return_pred)) + 
  scale_colour_gradient(low = "green", high = "blue") +
  geom_line()
```

```{r}
ggplot(signal, aes(x = date, y = close, colour = future_return_pred)) + 
  scale_colour_gradient(low = "green", high = "blue") +
  geom_line() + 
  scale_y_log10()
```

The following plots show the signal translated into long and flat positions using a 0.50 probability threshold for going long.

```{r}
ggplot(signal, aes(x = date, y = close, colour = as.numeric(signal))) + 
  scale_colour_gradient(low = "green", high = "blue") +
  geom_line()
```

```{r}
ggplot(signal, aes(x = date, y = close, colour = as.numeric(signal))) + 
  scale_colour_gradient(low = "green", high = "blue") +
  geom_line() + 
  scale_y_log10()
```

The following plots show the equity curve of the model compared to a buy-and-hold return.

```{r}
ggplot(signal, aes(x = date)) + 
  geom_line(aes(y = return_model), colour = "red") + 
  geom_line(aes(y = return_buyhold), colour = "blue")
```

```{r}
ggplot(signal, aes(x = date)) + 
  geom_line(aes(y = return_model), colour = "red") + 
  geom_line(aes(y = return_buyhold), colour = "blue") + 
  scale_y_log10()
```

The following plots show the model return minus the buy-and-hold return.

```{r}
ggplot(signal, aes(x = date)) + 
  geom_line(aes(y = return_model - return_buyhold), colour = "red")
```

```{r}
ggplot(signal, aes(x = date)) + 
  geom_line(aes(y = return_model - return_buyhold + 1), colour = "red") + 
  scale_y_log10()
```

## 7.2 Descriptive Statistics

### 7.2.1 Confusion Matrix

```{r}
conf_mat(signal %>% mutate(signal = as.factor(signal)), truth = "future_return_sign", estimate = 'signal')
```

### 7.2.2 Accuracy

```{r}
accuracy(signal %>% mutate(signal = as.factor(signal)), truth = 'future_return_sign', estimate = 'signal')
```

### 7.2.3 Log Loss

```{r}
mn_log_loss(signal %>% mutate(signal = as.factor(signal)), truth = 'future_return_sign', estimate = 'future_return_pred')
```

### 7.2.4 Overall Return of Model vs Buy-and-Hold

```{r}
signal %>% select(return_model) %>% filter(row_number() == n())
```

```{r}
signal %>% select(return_buyhold) %>% filter(row_number() == n())
```

### 7.2.5 Mean Daily Return & SD

```{r}
mean(signal$close_change_1d * as.numeric(signal$signal))
```

```{r}
sd(signal$close_change_1d * as.numeric(signal$signal))
```

### 7.2.6 Annualised Return & SD

```{r}
(1 + mean(signal$close_change_1d * signal$signal))^365
```

```{r}
sd(signal$close_change_1d * signal$signal) * 365^0.5
```

### 7.2.7 Annualised Sharpe Ratio

```{r}
mean(signal$close_change_1d * as.numeric(signal$signal)) / 
  sd(signal$close_change_1d * as.numeric(signal$signal)) *
  365^0.5
```

### 7.2.8 Max Drawdown of Model vs Buy-and-Hold Return

```{r}
signal %>% 
  mutate(return_model_drawdown = -1 * (1 - return_model / cummax(return_model))) %>%
  select(return_model_drawdown) %>%
  min()
```

```{r}
signal %>% 
  mutate(return_buyhold_drawdown = -1 * (1 - return_buyhold / cummax(return_buyhold))) %>% 
  select(return_buyhold_drawdown) %>% 
  min()
```

### 7.2.9 Daily & Annualized Semideviation

```{r}
down <- signal %>%
  filter(close_change_1d * as.numeric(signal) < 0)

sd(down$close_change_1d * down$signal)
```

```{r}
sd(down$close_change_1d * down$signal) * 365^0.5
```

### 7.2.10 Annualised Sortino Ratio

```{r}
mean(signal$close_change_1d * signal$signal) / 
  sd(down$close_change_1d * down$signal) * 365^0.5
```

## 7.3 Most Recent Predictions

```{r}
signal %>% 
  select(date, close, future_return_pred) %>%
  tail(n = 30)
```

```{r}
temp <- signal %>%
  select(date, close, future_return_pred, signal) %>%
  filter(date >= '2017-08-01')

rmarkdown::paged_table(temp)
```

```{r}
ggplot(temp, aes(x = date)) + 
  geom_line(aes(y = close, colour = signal)) + 
  geom_point(aes(y = close, colour = signal)) + 
  scale_color_gradient(low = "red", high = "blue")
```

```{r}
ggplot(temp, aes(x = date)) + 
  geom_line(aes(y = close, colour = future_return_pred)) + 
  geom_point(aes(y = close, colour = future_return_pred)) + 
  scale_color_gradient2(low = "red", mid = "grey80", high = "blue", midpoint = 0.50)
```